A comprehensive framework to capture the arcana of neuroimaging analysis

### 2019-03-25 manuscript/manuscript.tex
removed relative links to graphics and code listings
final submission

\begin{figure*}
    \centering
    \includegraphics[width=0.55\textwidth]{../figures/example_study}
  \caption{Example study. Blue boxes represent input
data (filesets or fields) stored in a repository and green derivatives
from that data stored alongside the original data. Orange ovals are
pipelines that operate on data in the repository to derive the
derivatives. Arrows represent data flows, i.e., inputs and outputs to
pipelines}
\label{fig:example_study}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{../figures/simplified_arcana_uml}
  \caption{Unified Modelling Language (UML) diagram
of information flow in the Arcana framework. Boxes: Python classes,
blue=analysis-design, green=analysis-application. Arrows: orange=data,
magenta=workflow description, diamond=aggregated-in. Study classes
construct analysis pipelines, which are sent to the \emph{Processor} to
be processed. Input data is selected by \emph{Input} objects and
pulled to the compute environment to be processed along with existing
intermediate derivatives. After the derivatives are pushed back to the
repository.}
\label{fig:simple_uml}
\end{figure*}

@@

\begin{figure}
\lstinputlisting[basicstyle=\footnotesize,firstline=2, lastline=59]{../code/example_study.py}
\caption{Example data and parameter specifications. The data
specification specifies two input file-sets, `one' and `ten' and
ten derived file-sets that can be derived from them, at least
indirectly. Each derived data spec, specifies the name of the pipeline
constructor that creates the pipeline that derives them. Parameter
specifications specify a name and default value for free parameters of
the Study class}
\label{fig:data_spec}
\end{figure}

@@

\begin{figure}
\lstinputlisting[firstline=61, basicstyle=\footnotesize]{../code/example_study.py}
\caption{Example pipeline constructor method. Pipelines are
created using the \emph{pipeline} method of the Study class.
Pipeline objects are thin wrappers around Nipype Workflow objects to
in order manage the namespaces of the workflow's inputs, outputs and nodes. Every
pipeline constructor method should allow wildcard keyword arguments,
which are passed to the \emph{name\_maps} argument of the pipeline
initialisation. This allows pipeline constructors in sub and multi classes to map the
inputs and outputs of the pipeline onto different data specifications.}
\label{fig:pipeline_constructor}
\end{figure}

@@

\begin{figure*}
  \centering
    \includegraphics[width=0.775\textwidth]{../figures/example_multi_study}
  \caption{Example MultiStudy. Blue boxes represent input data
(filesets or fields) and green derivatives. Orange ovals
are pipelines. Blue and green arrows: pipeline inputs from study inputs
and derived data, respectively. Orange arrows: outputs of pipelines. Dashed
boxes represent data specifications in a sub-study that are present in
the global namespace and mapped into the sub-study space, and dotted
arrows the mappings. Sub-studies are linked by mapping the same data
spec in the global space onto data specifications the multiple sub-study
namespaces (e.g. \emph{Derived} \emph{1}, \emph{2} and \emph{3}).
There are no restrictions between mapping study input and derivative data
specifications: both input and derivative specifications can be mapped
onto input data or derivative specifications in sub-studies.
If a spec in the global spaced is mapped onto a derivative spec in the
sub-study space, then the pipeline that generates that derivative in the
sub-study will not run unless it generates other required derivatives
(e.g. \emph{Pipeline 3} in \emph{Sub-Study 2})}
\label{fig:example_multi_study}
\end{figure*}

@@

\begin{figure*}
	\centering
    \includegraphics[width=\textwidth]{../figures/full_arcana_uml}
  \caption{Detailed Unified Modelling Language (UML) diagram of
information flow in the Arcana framework. Boxes: Python classes
(blue=core, green=interchangeable modules, grey=example
specialisations). Arrows: orange=data, magenta=workflow description,
diamond=aggregated-in, triangle=subclass-of. Calling \emph{data(name)}
on a Study subclass constructs the requisite pipelines (as specified in
\emph{data\_specs}) to produce the requested data, and sends them
to the \emph{Processor} to be processed. Data is selected by
\emph{Input} objects, pulled to the compute environment to be
processed, and then the derivatives are pushed back to the repository.
Repositories can be of plain directories, or BIDS or XNAT repositories}
\label{fig:full_uml}
\end{figure*}

@@

\begin{figure}
\lstinputlisting[basicstyle=\footnotesize]{../code/example_application.py}
\caption{Example application of Study class to a dataset stored in an XNAT
repository. Once the Study object has been initialised potential
derivatives of the Study can be requested, and will be generated and
stored in the repository if already present.}
\label{fig:study_application}
\end{figure}

@@

\begin{figure*}
	\centering
    \includegraphics[width=0.55\textwidth]{../figures/dwi_study}
  \caption{Example diffusion-weighted MRI (DWI) study, which can derive
tensor metrics, fractional anisotropy (FA) and apparent diffusion
coefficient (ADC) as well as streamlines fibre tracking. Blue boxes:
acquired (input) data (filesets or fields). Green boxes: derivatives.
Orange ovals: pipelines. Blue and green arrows: acquired and derived
inputs to pipelines, respectively. Orange arrows: outputs of pipelines.
The DWI magnitude image is preprocessed for motion correction and EPI
distortions masked and bias corrected. From the bias corrected image two
branches of analysis can be performed using the same intermediate
derivatives: FA and ADC and/or streamlines fibre tracking.}
\label{fig:dwi_study}
\end{figure*}

@@

\begin{figure*}
	\centering
    \includegraphics[width=0.775\textwidth]{../figures/t2star_t1_study}
  \caption{Combined T2*/T1-weighted studies within \emph{ArcanaPaper}
  MultiStudy class, which can derive vein masks by combining Quantitative Susceptibility
Mapping (QSM) and Susceptible Weighted Imaging (SWI) contrasts with a
manual atlas. Blue boxes: input data (filesets or fields).
Green boxes: derivatives. Orange ovals: pipelines. Blue and green
arrows: inputs to pipelines from input and derived data, respectively. Orange
arrows: outputs of pipelines. Dashed boxes represent data specifications
in a sub-study that are present in the global namespace and mapped into
the sub-study space, and dotted arrows the mappings. The acquired
T1-weighted image is mapped to both the \emph{magnitude} spec of the
T1-weighted sub-study and the \emph{registration reference} spec of the
T2*-weighted sub-study. The nonlinear transformation from subject to
atlas space are mapped from the T1-weighted sub-study and combined with
the linear registration between T1-weighted and T2*-weighted images, QSM
and SWI images are combined to produce the composite-vein image. In this
instance, the SWI acquired from the scanner console is passed
as an input to the derived \emph{SWI} specification, overriding the
\emph{SWI} pipeline that would otherwise generate it (dashed oval).}
\label{fig:t2star_t1_study}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{../figures/veins}
  \caption{Composite vein image (\emph{third row}) constructed by
  combining susceptibility weighted imaging (SWI) (\emph{top row}),
  quantitative susceptibility mapping (QSM) (\emph{second row})
  and a vein atlas from manual tracings. A vein mask was then
  generated (\emph{bottom row}) from the composite
  vein image.
  \emph{left column}: axial slices. \emph{centre column}: coronal slices.
  \emph{right column}: sagittal slices.}
\label{fig:veins}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{../figures/fa_adc}
  \caption{Fractional Anisotropy (FA) \emph{(top row)} and
Apparent Diffusion Coefficient (ADC) (\emph{bottom rowI)} derived from
diffusion MRI data. \emph{Left column:} axial midline slices.
\emph{Middle column:} coronal midline slices. \emph{Left column}:
sagittal midline slices.}
\label{fig:fa_adc}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{../figures/tractography}
  \caption{Global tractography performed using the MRtrix
toolbox. Probabilistic streamlines generated with the iFOD2 algorithm
from fibre Orientation Distribution Function (fODF) estimated from
diffusion MRI datasets using Constrained Spherical Deconvolution (CSD).
Streamlines are colour-encoded by orientation: green=anterior-posterior,
blue=inferior-superior, red=left-right. \emph{Left panel:} axial midline
slice. \emph{Middle panel:} coronal midline slice. \emph{Left panel}:
sagittal midline slice.}
\label{fig:tractography}
\end{figure*}

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=0.55\textwidth]{example_study}
  \caption{Example study. Blue boxes represent input
data (filesets or fields) stored in a repository and green derivatives
from that data stored alongside the original data. Orange ovals are
pipelines that operate on data in the repository to derive the
derivatives. Arrows represent data flows, i.e., inputs and outputs to
pipelines}
\label{fig:example_study}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{simplified_arcana_uml}
  \caption{Unified Modelling Language (UML) diagram
of information flow in the Arcana framework. Boxes: Python classes,
blue=analysis-design, green=analysis-application. Arrows: orange=data,
magenta=workflow description, diamond=aggregated-in. Study classes
construct analysis pipelines, which are sent to the \emph{Processor} to
be processed. Input data is selected by \emph{Input} objects and
pulled to the compute environment to be processed along with existing
intermediate derivatives. After the derivatives are pushed back to the
repository.}
\label{fig:simple_uml}
\end{figure*}

@@

\begin{figure}
\lstinputlisting[language=Python,basicstyle=\footnotesize,firstline=2, lastline=59]{example_study.py.txt}
\caption{Example data and parameter specifications. The data
specification specifies two input file-sets, `one' and `ten' and
ten derived file-sets that can be derived from them, at least
indirectly. Each derived data spec, specifies the name of the pipeline
constructor that creates the pipeline that derives them. Parameter
specifications specify a name and default value for free parameters of
the Study class}
\label{fig:data_spec}
\end{figure}

@@

\begin{figure}
\lstinputlisting[language=Python, firstline=61, basicstyle=\footnotesize]{example_study.py.txt}
\caption{Example pipeline constructor method. Pipelines are
created using the \emph{pipeline} method of the Study class.
Pipeline objects are thin wrappers around Nipype Workflow objects to
in order manage the namespaces of the workflow's inputs, outputs and nodes. Every
pipeline constructor method should allow wildcard keyword arguments,
which are passed to the \emph{name\_maps} argument of the pipeline
initialisation. This allows pipeline constructors in sub and multi classes to map the
inputs and outputs of the pipeline onto different data specifications.}
\label{fig:pipeline_constructor}
\end{figure}

@@

\begin{figure*}
  \centering
    \includegraphics[width=0.775\textwidth]{example_multi_study}
  \caption{Example MultiStudy. Blue boxes represent input data
(filesets or fields) and green derivatives. Orange ovals
are pipelines. Blue and green arrows: pipeline inputs from study inputs
and derived data, respectively. Orange arrows: outputs of pipelines. Dashed
boxes represent data specifications in a sub-study that are present in
the global namespace and mapped into the sub-study space, and dotted
arrows the mappings. Sub-studies are linked by mapping the same data
spec in the global space onto data specifications the multiple sub-study
namespaces (e.g. \emph{Derived} \emph{1}, \emph{2} and \emph{3}).
There are no restrictions between mapping study input and derivative data
specifications: both input and derivative specifications can be mapped
onto input data or derivative specifications in sub-studies.
If a spec in the global spaced is mapped onto a derivative spec in the
sub-study space, then the pipeline that generates that derivative in the
sub-study will not run unless it generates other required derivatives
(e.g. \emph{Pipeline 3} in \emph{Sub-Study 2})}
\label{fig:example_multi_study}
\end{figure*}

@@

\begin{figure*}
	\centering
    \includegraphics[width=\textwidth]{full_arcana_uml}
  \caption{Detailed Unified Modelling Language (UML) diagram of
information flow in the Arcana framework. Boxes: Python classes
(blue=core, green=interchangeable modules, grey=example
specialisations). Arrows: orange=data, magenta=workflow description,
diamond=aggregated-in, triangle=subclass-of. Calling \emph{data(name)}
on a Study subclass constructs the requisite pipelines (as specified in
\emph{data\_specs}) to produce the requested data, and sends them
to the \emph{Processor} to be processed. Data is selected by
\emph{Input} objects, pulled to the compute environment to be
processed, and then the derivatives are pushed back to the repository.
Repositories can be of plain directories, or BIDS or XNAT repositories}
\label{fig:full_uml}
\end{figure*}

@@

\begin{figure}
\lstinputlisting[language=Python, basicstyle=\footnotesize]{example_application.py.txt}
\caption{Example application of Study class to a dataset stored in an XNAT
repository. Once the Study object has been initialised potential
derivatives of the Study can be requested, and will be generated and
stored in the repository if already present.}
\label{fig:study_application}
\end{figure}

@@

\begin{figure*}
	\centering
    \includegraphics[width=0.55\textwidth]{dwi_study}
  \caption{Example diffusion-weighted MRI (DWI) study, which can derive
tensor metrics, fractional anisotropy (FA) and apparent diffusion
coefficient (ADC) as well as streamlines fibre tracking. Blue boxes:
acquired (input) data (filesets or fields). Green boxes: derivatives.
Orange ovals: pipelines. Blue and green arrows: acquired and derived
inputs to pipelines, respectively. Orange arrows: outputs of pipelines.
The DWI magnitude image is preprocessed for motion correction and EPI
distortions masked and bias corrected. From the bias corrected image two
branches of analysis can be performed using the same intermediate
derivatives: FA and ADC and/or streamlines fibre tracking.}
\label{fig:dwi_study}
\end{figure*}

@@

\begin{figure*}
	\centering
    \includegraphics[width=0.775\textwidth]{t2star_t1_study}
  \caption{Combined T2*/T1-weighted studies within \emph{ArcanaPaper}
  MultiStudy class, which can derive vein masks by combining Quantitative Susceptibility
Mapping (QSM) and Susceptible Weighted Imaging (SWI) contrasts with a
manual atlas. Blue boxes: input data (filesets or fields).
Green boxes: derivatives. Orange ovals: pipelines. Blue and green
arrows: inputs to pipelines from input and derived data, respectively. Orange
arrows: outputs of pipelines. Dashed boxes represent data specifications
in a sub-study that are present in the global namespace and mapped into
the sub-study space, and dotted arrows the mappings. The acquired
T1-weighted image is mapped to both the \emph{magnitude} spec of the
T1-weighted sub-study and the \emph{registration reference} spec of the
T2*-weighted sub-study. The nonlinear transformation from subject to
atlas space are mapped from the T1-weighted sub-study and combined with
the linear registration between T1-weighted and T2*-weighted images, QSM
and SWI images are combined to produce the composite-vein image. In this
instance, the SWI acquired from the scanner console is passed
as an input to the derived \emph{SWI} specification, overriding the
\emph{SWI} pipeline that would otherwise generate it (dashed oval).}
\label{fig:t2star_t1_study}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{veins}
  \caption{Composite vein image (\emph{third row}) constructed by
  combining susceptibility weighted imaging (SWI) (\emph{top row}),
  quantitative susceptibility mapping (QSM) (\emph{second row})
  and a vein atlas from manual tracings. A vein mask was then
  generated (\emph{bottom row}) from the composite
  vein image.
  \emph{left column}: axial slices. \emph{centre column}: coronal slices.
  \emph{right column}: sagittal slices.}
\label{fig:veins}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fa_adc}
  \caption{Fractional Anisotropy (FA) \emph{(top row)} and
Apparent Diffusion Coefficient (ADC) (\emph{bottom rowI)} derived from
diffusion MRI data. \emph{Left column:} axial midline slices.
\emph{Middle column:} coronal midline slices. \emph{Left column}:
sagittal midline slices.}
\label{fig:fa_adc}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{tractography}
  \caption{Global tractography performed using the MRtrix
toolbox. Probabilistic streamlines generated with the iFOD2 algorithm
from fibre Orientation Distribution Function (fODF) estimated from
diffusion MRI datasets using Constrained Spherical Deconvolution (CSD).
Streamlines are colour-encoded by orientation: green=anterior-posterior,
blue=inferior-superior, red=left-right. \emph{Left panel:} axial midline
slice. \emph{Middle panel:} coronal midline slice. \emph{Left panel}:
sagittal midline slice.}
\label{fig:tractography}
\end{figure*}
